---
title: "LAScatalog_PE"
output: html_document
date: "2025-03-18"
---

## Setup
```{r setup, include=FALSE}
# Clear environment
rm(list = ls(globalenv()))

# Load packages
library(lidR)
library(mapview)
library(sf)
library(future)
library(stringr)
```

```{r}
# Set up threads
threads = 8
get_lidr_threads()
set_lidr_threads(threads)
```


## LAScatalog processing engine
### Create LAScatalog object
```{r}
ctg <- readLAScatalog("tiles_data/bretten01_leafon/las_output")
```

```{r}
# For LAScatalog object
crs(ctg) <- 25832
st_crs(ctg) <- 25832
st_bbox(ctg)
```

```{r}
summary(ctg)
plot(ctg, map= TRUE)
```

### Region of interset
```{r}
# Specify the filenames to remove
file_to_remove <- c("tiles_data/bretten01_leafon/las_output/476616.0870000124_5428970.0309999995_8.laz",
                    "tiles_data/bretten01_leafon/las_output/476616.08599996567_5429057.297000885_10.laz", 
                    "tiles_data/bretten01_leafon/las_output/476624.8219999969_5429157.286998749_11.laz", 
                    "tiles_data/bretten01_leafon/las_output/476616.0869998932_5429157.287998199_12.laz")

# Filter the catalog to exclude the specified file
ctg_filtered <- ctg[!ctg@data$filename %in% file_to_remove, ]
```

```{r}
plot(ctg_filtered, map = TRUE)
```

```{r}
# check if files have .lax
is.indexed(ctg_filtered)

# generate index files
lidR:::catalog_laxindex(ctg_filtered)
```


## Parallel processing
### Specify chunk size and buffer
```{r}
opt_chunk_size(ctg_filtered) <- 40
opt_chunk_buffer(ctg_filtered) <- 10

# Visualize and summarize the catalog chunks
plot(ctg_filtered, chunk = TRUE)
```

```{r}
#las_check(ctg_filtered, deep = TRUE)
summary(ctg_filtered)
```


## (Optional) Ground Classification 
```{r eval=FALSE, include=FALSE}
# Set the output directory
output_dir <- "tiles_data/bretten01_leafon/classified"
dir.create(output_dir, recursive = TRUE, showWarnings = FALSE)

# Dynamically generate output filenames using a template
opt_output_files(ctg_filtered) <- file.path(output_dir, "{XLEFT}_{YBOTTOM}_{ID}_classified")

# Classify the ground
classified_ctg <- classify_ground(ctg_filtered, pmf(ws = 5, th = 3))
```


## DTM 
```{r}
# Set the output directory and Generate output filenames using a template
output_dir <- "tiles_data/bretten01_leafon/dtm_0.1"
dir.create(output_dir, recursive = TRUE, showWarnings = FALSE)
opt_output_files(ctg_filtered) <- file.path(output_dir, "{XLEFT}_{YBOTTOM}_{ID}_dtm")

dtm <- rasterize_terrain(ctg_filtered, 0.1, knnidw())
```

## Height normalization
```{r}
output_dir <- "tiles_data/bretten01_leafon/norm_0.1"
dir.create(output_dir, recursive = TRUE, showWarnings = FALSE)
opt_output_files(ctg_filtered) <- file.path(output_dir, "{XLEFT}_{YBOTTOM}_{ID}_norm")

ctg_norm <- normalize_height(ctg_filtered, dtm)
```


## Import files from LAScatalog
```{r}
norm_input_dir <- "tiles_data/bretten01_leafon/norm_0.1"
ctg_norm <- readLAScatalog(norm_input_dir)
plot(ctg_norm, chunk=TRUE, map=TRUE)
```


## Local Maximum Filter 
```{r}
# Variable window definition
f <- function(x) {
  y <- numeric(length(x))
  y[x < 5] <- 1
  y[x >= 5 & x <= 10] <- 2
  y[x >= 10 & x <= 20] <- 3
  y[x >= 20 & x <=50 ] <- 4
  y[x > 50] <- 5
  return(y)
}

heights <- seq(-5,100,0.5)
ws <- f(heights)
plot(heights, ws, type = "l",  ylim = c(0,5), main = "Local Maximum Filter with Variable Window Size")
```


## Resolutions/algorithoms Exploration for tree tops detection
```{r}
# Point-to-raster 2 resolutions
chm_p2r_05 <- rasterize_canopy(ctg_norm, 0.5, p2r(subcircle = 0.2), pkg = "terra")
chm_p2r_1 <- rasterize_canopy(ctg_norm, 1, p2r(subcircle = 0.2), pkg = "terra")
```

```{r}
# Pitfree with and without subcircle tweak
chm_pitfree_05_1 <- rasterize_canopy(ctg_norm, 0.5, pitfree(), pkg = "terra")
chm_pitfree_05_2 <- rasterize_canopy(ctg_norm, 0.5, pitfree(subcircle = 0.2), pkg = "terra")
```

```{r}
# Post-processing median filter
chm_p2r_05_smoothed <- terra::focal(chm_p2r_05, w = kernel, fun = median, na.rm = TRUE)
chm_p2r_1_smoothed <- terra::focal(chm_p2r_1, w = kernel, fun = median, na.rm = TRUE)
```

```{r}
ttops_chm_p2r_05 <- locate_trees(chm_p2r_05, lmf(f))
ttops_chm_p2r_1 <- locate_trees(chm_p2r_1, lmf(f))
ttops_chm_pitfree_05_1 <- locate_trees(chm_pitfree_05_1, lmf(f))
ttops_chm_pitfree_05_2 <- locate_trees(chm_pitfree_05_2, lmf(f))
ttops_chm_p2r_05_smoothed <- locate_trees(chm_p2r_05_smoothed, lmf(f))
ttops_chm_p2r_1_smoothed <- locate_trees(chm_p2r_1_smoothed, lmf(f))
```


### CHM based Segmentation
```{r}
output_dir <- "tiles_data/bretten01_leafon/chm_1"
dir.create(output_dir, recursive = TRUE, showWarnings = FALSE)
opt_output_files(ctg_norm) <- file.path(output_dir, "{XLEFT}_{YBOTTOM}_{ID}_chm")

kernel <- matrix(1,3,3)
chm_p2r_1 <- rasterize_canopy(ctg_norm, 1, p2r(subcircle = 0.2), pkg = "terra")
```

```{r}
col <- height.colors(50)
plot(chm_p2r_1, main = "CHM P2R 1", col = col)
```


```{r}
chm_p2r_1_smoothed <- terra::focal(chm_p2r_1, w = kernel, fun = median, na.rm = TRUE)
ttops_chm_p2r_1_smoothed <- locate_trees(chm_p2r_1_smoothed, lmf(f), uniqueness = "bitmerge")
```

```{r}
plot(chm_p2r_1_smoothed, main = "CHM P2R 1 smoothed", col = col) 
plot(sf::st_geometry(ttops_chm_p2r_1_smoothed), add = T, pch =3)
```

```{r}
output_dir <- "tiles_data/bretten01_leafon/seg_1"
dir.create(output_dir, recursive = TRUE, showWarnings = FALSE)
opt_output_files(ctg) <- file.path(output_dir, "{XLEFT}_{YBOTTOM}_{ID}_seg")
algo <- dalponte2016(chm_p2r_1_smoothed, ttops_chm_p2r_1_smoothed)
tree_segs <- segment_trees(ctg, algo) # segment point cloud
```



## Plotting
```{r}
dtm
plot(dtm, main = "Digital Terrain Model (DTM)", xlab = "Easting", ylab = "Northing", legend = TRUE)
```


```{r eval=FALSE, include=FALSE}
hmean <- pixel_metrics(ctg_norm, ~mean(Z), 1)
plot(hmean, col = height.colors(25))
```


```{r}
col <- height.colors(50)
plot(chm_p2r_05, main = "CHM P2R 0.5", col = col); plot(sf::st_geometry(ttops_chm_p2r_05), add = T, pch =3)
plot(chm_p2r_1, main = "CHM P2R 1", col = col); plot(sf::st_geometry(ttops_chm_p2r_1), add = T, pch = 3)
plot(chm_p2r_05_smoothed, main = "CHM P2R 0.5 smoothed", col = col); plot(sf::st_geometry(ttops_chm_p2r_05_smoothed), add = T, pch =3)
plot(chm_p2r_1_smoothed, main = "CHM P2R 1 smoothed", col = col); plot(sf::st_geometry(ttops_chm_p2r_1_smoothed), add = T, pch =3)
plot(chm_pitfree_05_1, main = "CHM PITFREE 1", col = col); plot(sf::st_geometry(ttops_chm_pitfree_05_1), add = T, pch =3)
plot(chm_pitfree_05_2, main = "CHM PITFREE 2", col = col); plot(sf::st_geometry(ttops_chm_pitfree_05_2), add = T, pch =3)

```


```{r}
opt_output_files(ctg_segmented) <- ""
lasplot <- clip_circle(ctg_segmented, 338500, 5238600, 40)
pol = crown_metrics(lasplot, NULL, geom = "convex")

plot(sf::st_geometry(pol), col = pastel.colors(250), axes = T)
plot(ctg, add = T)
```


```{r}

sum(ttops$treeID == 1)
#> [1] 9
plot(ttops["treeID"], cex = 0.3, pch = 19, nbreaks = 50)

ttops$treeID <- 1:nrow(ttops)
ttops <- locate_trees(ctg_norm, lmf(4), uniqueness = "bitmerge")
plot(ttops["treeID"], cex = 0.01, pch = 19)


plot(chm, col = height.colors(50))
```

```{r}
ttops_chm_p2r_1_smoothed
plot(ttops_chm_p2r_1_smoothed["Z"], cex = 0.001, pch = 19, pal = height.colors, nbreaks = 30)
```

```{r}
seg_input_dir <- "tiles_data/bretten01_leafon/seg_1/476624.8_5429057_4_seg.las"
seg <- readLAS(seg_input_dir)
plot(seg, bg = "white", size = 4, color = "treeID")
```


## Import all files for visualization

```{r}
library(lidR)

# 1. Read the directory containing segmented LAS files
seg_input_dir <- "tiles_data/bretten01_leafon/seg_1/"

# 2. Get paths to all LAS files
seg_files <- list.files(seg_input_dir, pattern = "\\.las$", full.names = TRUE)
if (length(seg_files) == 0) {
  stop("No LAS files found. Please check the directory path.")
}

# 3. Read all files and check for empty data or missing 'treeID'
seg_list <- lapply(seg_files, function(file) {
  las <- readLAS(file)
  if (is.empty(las)) {
    warning("File ", file, " is empty and will be skipped.")
    return(NULL)
  }
  if (!"treeID" %in% names(las@data)) {
    warning("File ", file, " is missing the 'treeID' field and will be skipped.")
    return(NULL)
  }
  return(las)
})

# Remove NULL elements from the list
seg_list <- Filter(Negate(is.null), seg_list)

# 4. Merge all non-empty LAS objects
if (length(seg_list) == 0) {
  stop("No valid LAS files to merge.")
}
seg_combined <- do.call(rbind, seg_list)

# 5. Visualize the merged result
plot(seg_combined, bg = "white", size = 4, color = "treeID")
```




