---
title: "Individual_Tree_detection"
output: html_document
date: "2025-03-18"
---

```{r setup, include=FALSE}
# Install required packages
require(lidR)
library(ggplot2)
library(rgl)

```

```{r}
threads = 8
get_lidr_threads()
set_lidr_threads(threads)
get_lidr_threads()
```

## Import and preprocessing Point cloud

```{r}
las_first_echo <- readLAS("tiles_data/bretten01_leafon/las_output/476724.82099998_5429057.2870000005_1.laz")
plot(las, color = "Z", bg = "white", axis = TRUE, legend = TRUE)
```

```{r}
offsets <- plot(las)
print(offsets)
#[1]  476638.1 5429059.4
```


```{r}
p1 <- c(476638.1, 5429059.4)
p2 <- c(477000, 5429070)
las_tr <- clip_transect(las, p1, p2, width = 5, xz = TRUE)
ggplot(payload(las_tr), aes(X,Z, color = Z)) + 
  geom_point(size = 0.5) + 
  coord_equal() + 
  theme_minimal() +
  scale_color_gradientn(colours = height.colors(50))
```

## Ground Classification

```{r}
las_ground <- classify_ground(las, algorithm = pmf(ws = 5, th = 3))
```

```{r}

# Check the unique classes in the classified LAS object
unique(las_ground$Classification)

```


## Ground rasterization

```{r}
plot(las_ground, color = "Classification", size = 3, bg = "white") 
```

```{r}
gnd <- filter_ground(las_ground)
plot(gnd, size = 3, bg = "white", color = "Classification")
```

```{r}
dtm <- rasterize_terrain(las_first_echo, 1, knnidw())
```


## Normalize Height

nlas <- las - dtm makes very simple subtractions without any interpolation. normalize_height(las, knnidw()) computes on-the-fly the interpolation of the ground points and estimates the exact elevation of every points. An hybrid method consists in the interpolation of the pixel of an already computed DTM.

```{r}
nlas <- las_first_echo - dtm
plot(nlas, size = 4, bg = "white")

```


```{r}
nlas_hybird <- normalize_height(las_first_echo, tin(), dtm = dtm)
```

```{r}

hist(filter_ground(nlas_hybird)$Z, breaks = seq(-60, 60, 0.1), main = "", xlab = "Elevation")

#las <- unnormalize_height(nlas)
```


## Remove noise
```{r}
# Filter the tree points ( class 0)
las_trees <- filter_poi(nlas_hybird, Classification == 0L)
plot(las_trees, color = "Z", bg = "white", axis = TRUE, legend = TRUE)
```


```{r}
# Filter tree points where return number is 1
trees_first_echo <- filter_poi(las_trees, ReturnNumber == 1L)

# Plot the filtered tree points with high density
plot(trees_first_echo, color = "Z", bg = "white", axis = TRUE, legend = TRUE)

```

```{r}
# Using IVF
noise <- classify_noise(las_trees, sor(15,7))

# Remove outliers using filter_poi()
tree_denoise <- filter_poi(noise, Classification != LASNOISE)

plot(tree_denoise, color = "Z", bg = "white", axis = TRUE, legend = TRUE)
```


### Rasterize Canopy


#### Parameters testing
```{r}
chm_01 <- rasterize_canopy(tree_denoise, res = 0.1, algorithm = p2r())
col <- height.colors(25)
plot(chm_01, col = col)

chm_005 <- rasterize_canopy(tree_denoise, res = 0.05, algorithm = p2r())
plot(chm_005, col = col)
```
One option to reduce the number of voids in the surface model is to replace every point in the point cloud with a disk of a known radius (e.g. 15 cm). This operation is meant to simulate the fact that the laser footprint is not a point, but rather a circular area. It is equivalent to computing the CHM from a densified point cloud in a way that tries to have a physical meaning.

```{r}
chm_01_cricle <- rasterize_canopy(tree_denoise, res = 0.1, algorithm = p2r(subcircle = 0.05))
plot(chm_01_cricle, col = col)

chm_01_tin <- rasterize_canopy(tree_denoise, res = 0.1, p2r(0.05, na.fill = tin()))
plot(chm_01_tin, col = col)

chm_dsmtin <- rasterize_canopy(tree_denoise, res = 0.1, algorithm = dsmtin())
plot(chm_dsmtin, col = col)

chm_005_cricle <- rasterize_canopy(tree_denoise, res = 0.05, algorithm = p2r(subcircle = 0.05))
plot(chm_005_cricle, col = col)

chm_005_tin <- rasterize_canopy(tree_denoise, res = 0.05, p2r(0.05, na.fill = tin()))
plot(chm_005_tin, col = col)

chm_005_dsmtin <- rasterize_canopy(tree_denoise, res = 0.05, algorithm = dsmtin())
plot(chm_005_dsmtin, col = col)

chm_01_cricle_01 <- rasterize_canopy(tree_denoise, res = 0.1, algorithm = p2r(subcircle = 0.1))
chm_005_cricle_01 <- rasterize_canopy(tree_denoise, res = 0.05, algorithm = p2r(subcircle = 0.1))
plot(chm_01_cricle_01, col = col)
plot(chm_005_cricle_01, col = col)

```


```{r}
chm <- rasterize_canopy(tree_denoise, res = 0.1, algorithm = dsmtin(max_edge = 0.4))
plot(chm, col = col)
```

```{r}
# The first layer is a regular triangulation
layer0 <- rasterize_canopy(tree_denoise, res = 0.1, algorithm = dsmtin())

# Triangulation of first return above 5 m
above10 <- filter_poi(tree_denoise, Z >= 5)
layer10 <- rasterize_canopy(above10, res = 0.1, algorithm = dsmtin(max_edge = 0.4))

# Triangulation of first return above 10 m
above20 <- filter_poi(above10, Z >= 10)
layer20 <- rasterize_canopy(above20, res = 0.1, algorithm = dsmtin(max_edge = 0.4))

# The final surface is a stack of the partial rasters
dsm <- layer0
dsm[] <- pmax(as.numeric(layer0[]), as.numeric(layer10[]), as.numeric(layer20[]), na.rm = T)

layers <- c(layer0, layer10, layer20, dsm)
names(layers) <- c("Base", "Layer10m", "Layer20m", "pitfree")
plot(layers, col = col)


```

```{r}
chm_pitfre_01 <- rasterize_canopy(tree_denoise, res = 0.1, pitfree(thresholds = c(0, 10, 20), max_edge = c(0, 1)))
plot(chm_pitfre_01, col = col)

chm_pitfre_01_circle <- rasterize_canopy(tree_denoise, res = 0.1, pitfree(subcircle = 0.1))
plot(chm_pitfre_01_circle, col = col)


```


##  Individual Tree Detection (ITD)

```{r}

fill.na <- function(x, i=5) { if (is.na(x)[i]) { return(mean(x, na.rm = TRUE)) } else { return(x[i]) }}
w <- matrix(1, 3, 3)

chm <- rasterize_canopy(tree_denoise, res = 0.1, algorithm = p2r(subcircle = 0.1), pkg = "terra")
filled <- terra::focal(chm, w, fun = fill.na)
smoothed <- terra::focal(chm, w, fun = mean, na.rm = TRUE)

chms <- c(chm, filled, smoothed)
names(chms) <- c("Base", "Filled", "Smoothed")
plot(chms, col = col)
```



```{r}

f <- function(x) {x * 0.1 + 3}
heights <- seq(0,30,5)
ws <- f(heights)
plot(heights, ws, type = "l", ylim = c(0,6))

```

```{r}
f <- function(x) {
  y <- numeric(length(x))
  y[x < 5] <- 1
  y[x >= 5 & x <= 10] <- 2
  y[x >= 10 & x <= 20] <- 3
  y[x > 20] <- 4  
  return(y)
}

heights <- seq(-5,40,0.5)
ws <- f(heights)
plot(heights, ws, type = "l",  ylim = c(0,5), main = "Local Maximum Filter with Variable Window Size")
```

```{r}
# Local Maximum Filter with variable windows size
ttops <- locate_trees(tree_denoise, lmf(f))

# Plot the CHM with the tree tops overlaid
plot(chm, col = height.colors(50), main = "Canopy Height Model with Tree Tops")
plot(sf::st_geometry(ttops), add = TRUE, pch = 3, col = "black")  # Add tree tops to the plot

```


### Apply rasterization
```{r}
# Point-to-raster 2 resolutions
chm_p2r_05 <- rasterize_canopy(tree_denoise, 0.5, p2r(subcircle = 0.2), pkg = "terra")
chm_p2r_1 <- rasterize_canopy(tree_denoise, 1, p2r(subcircle = 0.2), pkg = "terra")

# Pitfree with and without subcircle tweak
chm_pitfree_05_1 <- rasterize_canopy(tree_denoise, 0.5, pitfree(), pkg = "terra")
chm_pitfree_05_2 <- rasterize_canopy(tree_denoise, 0.5, pitfree(subcircle = 0.2), pkg = "terra")

# Post-processing median filter
kernel <- matrix(1,3,3)
chm_p2r_05_smoothed <- terra::focal(chm_p2r_05, w = kernel, fun = median, na.rm = TRUE)
chm_p2r_1_smoothed <- terra::focal(chm_p2r_1, w = kernel, fun = median, na.rm = TRUE)
```


### Tree top detection

```{r}
ttops_chm_p2r_05 <- locate_trees(chm_p2r_05, lmf(f))
ttops_chm_p2r_1 <- locate_trees(chm_p2r_1, lmf(f))
ttops_chm_pitfree_05_1 <- locate_trees(chm_pitfree_05_1, lmf(f))
ttops_chm_pitfree_05_2 <- locate_trees(chm_pitfree_05_2, lmf(f))
ttops_chm_p2r_05_smoothed <- locate_trees(chm_p2r_05_smoothed, lmf(f))
ttops_chm_p2r_1_smoothed <- locate_trees(chm_p2r_1_smoothed, lmf(f))
```


```{r}
par(mfrow=c(3,2))
col <- height.colors(50)
plot(chm_p2r_05, main = "CHM P2R 0.5", col = col); plot(sf::st_geometry(ttops_chm_p2r_05), add = T, pch =3)
plot(chm_p2r_1, main = "CHM P2R 1", col = col); plot(sf::st_geometry(ttops_chm_p2r_1), add = T, pch = 3)
plot(chm_p2r_05_smoothed, main = "CHM P2R 0.5 smoothed", col = col); plot(sf::st_geometry(ttops_chm_p2r_05_smoothed), add = T, pch =3)
plot(chm_p2r_1_smoothed, main = "CHM P2R 1 smoothed", col = col); plot(sf::st_geometry(ttops_chm_p2r_1_smoothed), add = T, pch =3)
plot(chm_pitfree_05_1, main = "CHM PITFREE 1", col = col); plot(sf::st_geometry(ttops_chm_pitfree_05_1), add = T, pch =3)
plot(chm_pitfree_05_2, main = "CHM PITFREE 2", col = col); plot(sf::st_geometry(ttops_chm_pitfree_05_2), add = T, pch =3)

```




### CHM based Segmentation
```{r}
algo <- dalponte2016(chm_p2r_05_smoothed, ttops_chm_p2r_05_smoothed)
tree_segs <- segment_trees(las, algo) # segment point cloud
```



```{r}
plot(tree_segs, bg = "white", size = 4, color = "treeID") # visualize trees
```

```{r}

tree20 <- filter_poi(tree_segs, treeID == 20)
plot(tree20, size = 3, bg = "white")

```


```{r}
# Point Cloud based segmentation -- slow
algo2 <- li2012()
tree_segs2 <- segment_trees(las, algo2, attribute = "IDli")
plot(las, add = x + c(100,0), bg = "white", size = 4, color = "IDli", colorPalette = pastel.colors(200))
```




## Extract crown metrics


```{r}
crowns <- crown_metrics(tree_segs, func = .stdtreemetrics, geom = "convex")
plot(crowns["convhull_area"], main = "Crown area (convex hull)")
```

```{r}
metrics <- crown_metrics(tree_segs, ~list(z_max = max(Z), z_mean = mean(Z))) # calculate tree metrics
head(metrics)

```


```{r}
plot(metrics["z_max"], pal = hcl.colors, pch = 19) # plot using z_max


```

```{r}

custom_crown_metrics <- function(z, i) { # user-defined function
  metrics <- list(
     z_max = max(z),   # max height
     z_sd = sd(z),     # vertical variability of points
     i_mean = mean(i), # mean intensity
     i_max  = max(i)   # max intensity
   )
   return(metrics) # output
}

ccm = ~custom_crown_metrics(z = Z, i = Intensity)
```


```{r}
metrics <- crown_metrics(tree_segs, func = ccm, geom = "convex")
plot(metrics["z_max"], pal = hcl.colors)

```

```{r}
metrics <- crown_metrics(tree_segs, func = ccm, geom = "concave")
plot(metrics["z_max"], pal = hcl.colors)
```


```{r}

metrics <- crown_metrics(tree_segs, ~list(imean = mean(Intensity))) # calculate tree intensity metrics

# Step 2: Filter trees based on intensity (e.g., intensity > 80)
high_intensity_trees <- metrics[metrics$imean > 80, ]

# Assuming 'high_intensity_trees' contains the filtered trees with intensity > 80
plot(high_intensity_trees, pdal = "imean", main = "High Intensity Trees")

```

